2H qSIP Calculations

### 2021-09-06 qSIP_calculations.Rmd
u_d versus rel_amount plot

```{r, message=FALSE, warning=FALSE}
library(tidyverse) # general data wrangling and plotting
library(isoreader) # reading the raw data files
library(isoprocessor) # processing the data
library(isotopia) # isotope utilities
library(readxl) # reading excel docs
library(ggsci) # some nice color palettes
library(ggborderline) # nice looking ggplot lines
library(ggrepel)
```

@@

## Generation Time Calculations

@@

peaks_mapped_active <- peaks_mapped %>% 
  filter(str_detect(compound, "STD", negate = TRUE))
```

```{r}
gen_calc_data <- peaks_mapped_active %>% 
  select(c(Analysis, compound, d2H, at2H, sample_id, inc_time_d, terrain, f_start)) %>% 
  mutate(at2H_label = 0.005) %>% 
  mutate(f_label = 0.005, # label is 0.5 at%
         a = 0.5,
         #R = (d2H/1000 +1) * ref_ratio,
         t = inc_time_d,
         #f_t = R / (1 + R) %>%  # 2F at time t where t = inc_time_d
         f_t = at2H) %>% 
  group_by(compound) %>% 
  mutate(
    t0 = 0,
    d_ft = as.numeric(f_t - f_start), # change in isotopic enrichment from start to time t
    d_t = t - t0, # change in time from start to time t (days)
    u_d = gencalc(a = a, FT = f_t, F0 = f_start, FL = f_label, t = d_t) # turnover (µ) in days^-1
  )

gen_calc_data
```

@@

```{r "shaded chromatogram", fig.width=8, fig.height=8}
peaks_mapped_active %>% ggplot(
  aes(x = rt, y = amp2, fill = as.numeric(at2H), label = compound)) +
  geom_col(aes(width = rt_end - rt_start)) +
  scale_fill_viridis_c() +
  facet_wrap(~sample_id, ncol = 1) +
  geom_label_repel(size = 2, box.padding = 1) +
  coord_cartesian(ylim = c(0, 1300)) +
  labs(fill = "2F",
         x = "Retention Time (s)",
         y = "Amplitude (mV)")
```

>>>>

```{r, message=FALSE, warning=FALSE}
library(tidyverse) # general data wrangling and plotting
library(isoreader) # reading the raw data files
library(isoprocessor) # processing the data
library(isotopia) # isotope utilities
library(readxl) # reading excel docs
library(ggsci) # some nice color palettes
library(ggborderline) # nice looking ggplot lines
library(ggrepel)
library(latex2exp)
library(ggdist)

source(file.path("libs", "visualization.R"))
source(file.path("libs", "chromleon.R"))
`%nin%` = Negate(`%in%`)
```

@@

# Generation Time Calculations

@@

peaks_mapped <- peaks_mapped %>% 
  mutate(rel_amount = area2 / sum(area2[compound != "IS"]))
```

## Remove standards from gencalcs
```{r}
peaks_mapped_active <- peaks_mapped %>% 
  filter(str_detect(compound, "STD", negate = TRUE))
```

```{r}
gen_calc_data <- peaks_mapped_active %>% 
  select(c(Analysis, compound, d2H, at2H, sample_id, inc_time_d, inc_time_d_str, terrain, f_start, rel_amount)) %>% 
  mutate(
    at2H_label = 0.005,
    f_label = 0.5, # label is 0.5 at%
    a = 0.5, # assimilation efficiency
    t0 = 0, # start time
    d_t = inc_time_d - t0, # change in time
    f_t = at2H, # restate 2F at time t
    u_d = gencalc(a = 0.5, # Assimilation efficiency
                  FT = f_t, # 2F at time t (at%)
                  F0 = f_start, # 2F at t0 (at%)
                  FL = f_label, # 2F of label (at%)
                  t = d_t), # change in time
         # Returns turnover (µ) in days^-1
         gen_d = log(2) / u_d # calculate generation time
  )
```

@@

## Chromatogram scaled by 2F

```{r "shaded chromatogram", fig.width=8, fig.height=8}
peaks_mapped_active %>% ggplot(
  aes(x = rt, y = amp2, fill = as.numeric(at2H), label = compound)) +
  geom_col(aes(width = rt_end - rt_start)) +
  scale_fill_viridis_c(option = "magma") +
  facet_wrap(~inc_time_d_str, ncol = 1) +
  geom_text_repel(size = 2, box.padding = 0.5, nudge_y = 150) +
  theme_classic() +
  coord_cartesian(ylim = c(0, 1300)) +
  labs(fill = TeX("$^2F$"),
         x = "Retention Time (s)",
         y = "Amplitude (mV)") +
  theme(strip.background = element_blank())
```

```{r}
gen_calc_data %>% filter(inc_time_d != 0) %>% 
  ggplot(
  aes(x = gen_d, 
      y = inc_time_d_str, 
      color = inc_time_d_str, 
      fill = inc_time_d_str)) +
  stat_slab(height = 0.3, alpha = 0.7) +
  stat_pointinterval(color = "black") +
  geom_point(
    color = "black",
    shape = 124,
    size = 3,
    alpha = 1,
    position = position_nudge(y = .03)) +
  geom_text_repel(aes(label = compound), 
                  color = "black", 
                  nudge_y = -0.3, 
                  size = 3) +
  scale_x_log10() +
  annotation_logticks(sides = "b") +
  scale_fill_npg() +
  scale_color_npg() +
  theme_classic() +
  labs(x = "Generation time (days)",
       y = "Incubation time",
       title = "Generation times of soil bacteria",
       subtitle = "Lipid-specific generation times of bacteria collected from Alpine Tundra",
       caption = "Soils collected from Niwot Ridge LTER") +
  theme(legend.position = "none",
        axis.ticks.y = element_blank())
```

# GC-FID Data

## Import GC-FID Data
```{r}
root <- file.path("data", "gc_fid_data_2020")
file_list = list.files(path = root)
file_list <- file_list %>% 
  str_subset("blank", negate = TRUE) %>% 
  str_subset("Blank", negate = TRUE) %>% 
  str_subset("GCMS-STD", negate = TRUE) %>% 
  str_subset("BLK", negate = TRUE) # remove procedural blanks
sample_names <- file_list %>% 
  str_replace("FID.....", "") %>% 
  str_replace("_100ul-nhex.xls", "") %>% 
  str_replace("_100ul.xls", "")

all_data <-
  tibble(
    sample = sample_names,
    file = file_list,
    data = map(file.path(root, file), read_chromeleon_export),
    injection_details = map(data, "injection_details"),
    integration_results = map(data, "integration_results")
  ) %>% select(-data)
```

## Unnest integration details

```{r}
results <- all_data %>% 
  select(-injection_details) %>% 
  unnest(integration_results) %>% 
  filter(nchar(`Peak Name`) > 0) %>% 
  # major compound groups
  mutate(
    group = `Peak Name` %>% str_remove("\\d+-OH") %>% parse_number() %>% abs() %>% {paste0("C", .)} %>% factor()
  )
```

```{r}
# Combine triplicate samples into one
results_trip <- results %>% 
  # Remove standards and fractions
  filter(sample != "FAME37.1") %>% 
  filter(sample != "TCBAME-2") %>% 
  filter(!str_detect(sample, "F1")) %>% 
  filter(!str_detect(sample, "F2")) %>% 
  filter(!str_detect(sample, "F3")) %>% 
  filter(!str_detect(sample, "cr")) %>%
  mutate(sample_rep = str_remove(sample, "[xyz]")) %>% 
  group_by(sample_rep, `Peak Name`) %>%
  # Generate summary stats
  summarize(rel_area_max = max(`Relative Area`, na.rm=TRUE),
            rel_area_min = min(`Relative Area`, na.rm=TRUE),
            rel_height_max = max(`Relative Height`, na.rm=TRUE),
            rel_height_min = min(`Relative Height`, na.rm=TRUE),
            `Retention Time` = mean(`Retention Time`, na.rm=TRUE),
            `Relative Area` = mean(`Relative Area`, na.rm=TRUE),
            `Relative Height` = mean(`Relative Height`, na.rm=TRUE)) %>% 
  # Add some basic metadata
  mutate(inc_time = case_when(
    str_detect(sample_rep, "...0") ~ 0,
    str_detect(sample_rep, "...3") ~ 3,
    str_detect(sample_rep, "...7") ~ 7),
    site = case_when(
      str_detect(sample_rep, "c") ~ "Conifer",
      str_detect(sample_rep, "t") ~ "Tundra",
      str_detect(sample_rep, "g") ~ "Grassland",
      str_detect(sample_rep, "m") ~ "Meadow")
  )
```

## PLFA Summary Plot
```{r}
p_72hr_FAs <- results_trip %>% filter(inc_time == "3") %>% 
  # Remove Standard Peaks
  #filter(str_detect(`Peak Name`, c("PAIBE", "21:0 (STD)", "23:0 (STD)"), negate = TRUE)) %>% 
  filter(`Peak Name` %nin% c("PAIBE", "21:0 (STD)", "23:0 (STD)")) %>% 
  ggplot(aes(x = `Retention Time`, 
             y = `Relative Area`, 
             color = sample_rep, 
             fill = sample_rep, 
             label = `Peak Name`)) +
  geom_col(width = 0.08) +
  geom_errorbar(aes(ymin=rel_area_min, 
                    ymax = rel_area_max), 
                color = "black", 
                width = 0.05) +
  facet_wrap(~sample_rep, 
             ncol = 1, 
             scales = "free_y") +
  geom_label_repel(fill = "black", 
                   color = "white", 
                   alpha = 0.7, 
                   size = 3, 
                   nudge_y = 2, 
                   segment.color = "black") +
  ggtitle("PLFA Profile: 72hr") +
  theme_bw() +
  scale_color_jama() +
  scale_fill_jama()
p_72hr_FAs
```

### IRMS relabund
```{r}
peaks_mapped_active %>% ggplot(aes(x = rt, y = rel_amount, color = inc_time_d_str)) +
  geom_col() +
  facet_wrap(~inc_time_d_str, ncol = 1)
```

### Relabund vs. activity
```{r}
gen_calc_data %>% filter(inc_time_d != 0) %>% 
  ggplot(
  aes(x = rel_amount, y = u_d, color = inc_time_d_str, label = compound)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE) +
  geom_text_repel(size = 3) +
  scale_color_npg()
```

## Merge IRMS and FID data

### PICK UP HERE: how do we match GC-FID compound data with IRMS compound data?

```{r}
results_trip_matched <- results_trip %>% 
  filter(`Peak Name` %in% gen_calc_data$compound)

matched_idx <- match(gen_calc_data$compound, results_trip_matched$`Peak Name`)
results_trip_matched$`Peak Name`[matched_idx]

combined_irms_fid <- gen_calc_data %>%
  group_by(terrain) %>% 
  mutate(relative_area = results_trip$`Relative Area`[""])
```
